---
title: "Make Climate GIFs"
author: "Ryan Peek"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_notebook:
    highlight: pygments
    theme: readable
---

# Making Looped GIFs of Climate Data

The goal of this is to demonstrate how to take snow or LANDSAT data and loop through to create GIFs of these plots through time.

## First Get Data

```{r libraries}

library(sharpshootR)  # CDEC.snow.courses, CDECquery, CDECsnowQuery
library(ggplot2)
library(dplyr)
library(viridis)
library(scales)
library(cowplot)
library(stringr)
library(readr)

```

```{r getsnowdata, eval=F, echo=T}

data(CDEC.snow.courses) # load CDEC data
snw<-CDEC.snow.courses
head(snw)
snw$id<-as.factor(snw$id)
snw$latitude<-as.numeric(snw$latitude)
snw$longitude<-as.numeric(snw$longitude)*-1
snw$apr1avg_in<-snw$april.1.Avg.inches
snw<-select(snw, course_number, id, elev_feet:longitude,apr1avg_in)
str(snw)

# CDEC snow data, historical to now for all Stations
snowhist<-read_csv("../data/snow/cdec_all_snow.csv")
head(snowhist)
snowhist$Datetime<-lubridate::mdy(snowhist$Datetime)
snowhist$Snow<-as.numeric(snowhist$Snow)
snowhist<-select(snowhist, Snow:Measure,Datetime:Month)
dfsnw<-left_join(snowhist, snw, by = c("Station"="id"))
dfsnw$Station<-as.factor(dfsnw$Station)
dfsnw$Measure<-as.factor(dfsnw$Measure)
summary(dfsnw)

snw.dat <- dfsnw %>% 
  filter(!is.na(Snow), 
         !is.na(latitude),
         !is.na(apr1avg_in))
summary(snw.dat)

write_rds(snw.dat, path="../data_output/cdec_snow_hist_all.rds")

```

Once data has been downloaded and joined with the Station ID, we then need to merge with the **Meadow** dataset. To do this requires calculating the nearest distance between spatial points using the `rgeos` package.

```{r joinmdws w snow stations}

snw.dat<-read_rds("../data_output/cdec_snow_hist_all.rds")

# filter to single measurement type
snw.df<-filter(snw.dat, Measure=="SNOW WC (INCHES)", Year>1983, Month>=1, Month<7) %>%
  group_by(course_number,Station, Year,latitude,longitude) %>% 
  summarize(snow_cm=2.54*mean(Snow,na.rm=T),
            snow_in=mean(Snow,na.rm=T)) %>% 
  as.data.frame()

head(snw.df)
table(snw.df$Year)

# Now use these packages to make spatial and calculate nearest distances
library(sp)
library(rgeos)

# make a spatial df w distinct station IDs
snw.stations <- snw.df %>% distinct(Station, .keep_all = T) %>% select(course_number:Station, latitude:snow_in) %>% as.data.frame
snw.sp <- snw.df %>% distinct(Station, .keep_all = T) %>% as.data.frame
coordinates(snw.sp) <- ~longitude+latitude # make spatial
class(snw.sp) # check type of spatial data frame

# calculate distances between each of snow station points
d <- gDistance(snw.sp,byid = T)
str(d)

# calculate the second shortest distance (closest distance is of point to itself, therefore use second shortest)
min.d <- apply(d, 1, function(x) order(x, decreasing=F)[2])
str(min.d)

# make new dataframe
newdata <- cbind(snw.stations, snw.stations[min.d,], apply(d, 1, function(x) sort(x, decreasing=F)[2]))

head(newdata)
colnames(newdata) <- c(colnames(snw.stations), 'neighbor_course','n_id', 'n.lat', 'n.long', 'n.snow_cm', 'n.snow_in', 'distance')
h(newdata)

# NOW NEED TO MERGE WITH MEADOWS
mod.hgm<-read_rds(path="../data_output/hgm_gee_mdws_dat.rds")

dfMDW<-mod.hgm %>% dplyr::distinct(UCDID, .keep_all = T) %>% 
  select(PLOT, UCDID, LAT_DD, LONG_DD, HUC12)
names(dfMDW)[3:4]<-c('latitude','longitude')
names(dfMDW)

# df2 = snow
dfSNW<-select(snw.stations, course_number:longitude)
head(dfSNW)

# promote the input lists to SpatialPointsDataFrames
coordinates(dfMDW) <- c("longitude", "latitude")
coordinates(dfSNW) <- c("longitude", "latitude")             

#  Define these vectors, used in the loop.
closestSiteVec <- vector(mode = "numeric",length = nrow(dfMDW))
minDistVec     <- vector(mode = "numeric",length = nrow(dfMDW))

# Get the vector index of the data closest to each Locations of interest.
# spDistsN1 usage: spDistsN1(pointList, pointToMatch, longlat)
# where:  pointList   : List of candidate points, pointToMatch: Single point for which we seek the closest point in pointList.
# longlat     : TRUE  computes Great Circle distance in km, FALSE computes Euclidean distance in units of input geographic coordinates

# run loop                       
for (i in 1 : nrow(dfMDW))
{
  distVec <- spDistsN1(dfSNW,dfMDW[i,],longlat = TRUE)
  minDistVec[i] <- min(distVec)
  closestSiteVec[i] <- which.min(distVec)
}

# make dataframe
PointAssign <- as.data.frame(dfSNW[closestSiteVec,c(1:2)])
mdw_snow.df = data.frame(coordinates(dfMDW),dfMDW$PLOT,dfMDW$UCDID,dfMDW$HUC12,
                        closestSiteVec,minDistVec,PointAssign)
# Update column names 
names(mdw_snow.df) <- c("long","lat","PLOT","UCDID","HUC12","closestSNOTEL_ID", "Distance","course_number", "Station","snw_long","snw_lat")

head(mdw_snow.df)

# write out
write_csv(mdw_snow.df,path="../data_output/cdec_snw_mdw_merge.csv")
write_rds(mdw_snow.df,path = "../data_output/cdec_snw_mdw_merge.rds")
save(mdw_snow.df,file = "../data_output/cdec_snw_mdw_merge.rda")

rm(d,CDEC.snow.courses,PointAssign, dfMDW, dfSNW, closestSiteVec, distVec, min.d, minDistVec, snow.sp, snw.sp)

```

## Plot A Google Map with Joined Data

```{r MakeBasePlot, fig.align="center"}

## Load Data
cdec.snow.mdw<-read_rds("../data_output/cdec_snw_mdw_merge.rds")
snow.dat<-read_rds("../data_output/cdec_snow_hist_all.rds")
#mod.hgm<-read_rds(path="../data_output/hgm_gee_mdws_dat.rds") # full model dataset
#load("../data_output/model_datasets_Aug_max.rda")

# filter to single measurement type & join with cdec.snow.mdw
snow.df<-filter(snow.dat, Measure=="SNOW WC (INCHES)", Month>=1) %>%
  group_by(course_number,Station, Year,latitude,longitude) %>% 
  summarize(snow_cm=2.54*mean(Snow,na.rm=T),
            snow_in=mean(Snow,na.rm=T)) %>% 
  as.data.frame()

snow.hgm.df<-inner_join(snow.df, cdec.snow.mdw, by="Station") %>% 
  rename("course_number"=course_number.x) %>% 
  select(-course_number.y)
  
summary(snow.hgm.df)
table(snow.hgm.df$Year) # Nearly every meadow has data....could plot for each Month if necessary

## Set the SNOW Palettes
library(RColorBrewer)
breaks<-(c(0,25,50,75,100,125,150,175,200)) # for color scale
palette<-c("darkred","red4","red2","lightcoral","lightpink","limegreen","green2","forestgreen")
paletteBlu<-brewer.pal(9,name="Blues")
paletteGnBu<-brewer.pal(9,name="GnBu")
paletteRdBu<-brewer.pal(9,name="RdBu")


library(googleVis)
library(ggmap)

# Get the basemap
ca <- get_map(
  #location='california',
  c(lon=-120,lat=38.7),
  zoom=7,crop=T,
  scale="auto",color="bw",source="google",
  maptype="terrain") # can change to terrain

gg <- ggmap(ca, extent='panel',padding=0) 

# call the basemap, also: extent="panel",fullpage=TRUE
#gg

# TEST PLOT (SIMPLE)
# ggCA <- gg + geom_point(data=snow.hgm.df[snow.hgm.df$Year==1984,], 
#                         aes(x=longitude, y=latitude, size=snow_cm),
#                         show.legend=T, pch=21, color="gray30",
#                         fill="blue2", alpha=0.6)+
#   theme_bw() + facet_wrap(~Year, ncol=1)
# print(ggCA) 

# WITH PALETTE
ggSnw <- gg + geom_point(data=snow.hgm.df[snow.hgm.df$Year==1998,], 
                        aes(x=longitude, y=latitude,size=snow_cm*1.2, fill=snow_cm),
                        show.legend=F, pch=21)+ 
  theme_bw() + ylab("Latitude") + xlab("Longitude") +
  theme(legend.position=c(1,1),legend.justification=c(1,1),
        legend.direction="vertical",legend.text=element_text(size=8),
        legend.title=element_text(size=8, face="bold"),
        legend.box="horizontal", panel.background = element_blank(),
        legend.box.just = c("top"), 
        legend.background = element_rect(fill=alpha('white', 0.6), colour = "gray30")) +
  scale_fill_gradientn("Snow Depth (cm)",
                       colours=paletteRdBu,
                       breaks=breaks,values=breaks, rescaler = function(x, ...) x,
                       oob = identity,limits=range(breaks), space="Lab")+
  facet_wrap(~Year, ncol = 1)

print(ggSnw)

```

## NOW SIMPLE LOOP!

Use a loop to create PNGs.

```{r cdec loop, eval=F, echo=T}

for (i in c(1984:max(snow.hgm.df$Year))){
  gg + geom_point(data=snow.hgm.df[snow.hgm.df$Year==i,], 
                        aes(x=longitude, y=latitude, size=snow_cm*1.2, fill=snow_cm),
                  show.legend=F, pch=21, color="gray30")+ 
    theme_bw() + ylab("Latitude") + xlab("Longitude") +
    theme(axis.text.x = element_text(angle = 60, vjust=0.15, size=8),
          legend.position=c(1,1),legend.justification=c(1,1),
          legend.direction="vertical",legend.text=element_text(size=8),
          legend.title=element_text(size=8, face="bold"),
          legend.box="horizontal", panel.background = element_blank(),
          legend.box.just = c("top"), 
          legend.background = element_rect(fill=alpha('white', 0.6), colour = "gray30")) +
    scale_fill_gradientn("Snow Depth (cm)",
                         colours=brewer.pal(9,name="RdBu"),
                         limits=range(breaks), breaks=breaks)+
    # scale_fill_gradientn("Snow Depth (cm)",colours=paletteRdBu,
    #                     breaks=breaks,values=breaks, rescaler = function(x, ...) x,
    #                     oob = identity,limits=range(breaks), space="Lab")+
    facet_wrap(~Year, ncol = 1)
    print(paste0("saving plot ", i))
    ggsave(filename = paste0("../fig_output/cdec/hgm_snowcdec_",i,".png"),
           width=8, height=8, units = "in", dpi=150)
}


```

## Now Loop + Animate!

This uses the `purrr` package to map over a ggplot function and build pngs. 

```{r cdec snow animation}

library(purrr)
library(magick)
library(RColorBrewer)
library(googleVis)
library(ggmap)

# Get the basemap
ca <- get_map(
  #location='california',
  c(lon=-120,lat=38.7),
  zoom=7,crop=T,
  scale="auto",color="bw",source="google",
  maptype="terrain") # can change to terrain

gg <- ggmap(ca, extent='panel',padding=0) # call the basemap, also: extent="panel",fullpage=TRUE
#gg

cdec_map <- function(Yr){
  
  gg + geom_point(data=snow.hgm.df[snow.hgm.df$Year==Yr,], 
                        aes(x=longitude, y=latitude, size=snow_cm*1.2, fill=snow_cm),
                  show.legend=F, pch=21, color="gray30")+ 
    theme_bw() + ylab("Latitude") + xlab("Longitude") +
    theme(axis.text.x = element_text(angle = 60, vjust=0.15, size=8),
          legend.position=c(1,1),legend.justification=c(1,1),
          legend.direction="vertical",legend.text=element_text(size=8),
          legend.title=element_text(size=8, face="bold"),
          legend.box="horizontal", panel.background = element_blank(),
          legend.box.just = c("top"), 
          legend.background = element_rect(fill=alpha('white', 0.6), colour = "gray30")) +
    scale_fill_gradientn("Snow Depth (cm)",
                         colours=brewer.pal(8,name="RdBu"), 
                         limits=range(breaks), breaks=breaks)+
    # scale_fill_gradientn("Snow Depth (cm)",colours=paletteRdBu,
    #                     breaks=breaks,values=breaks, rescaler = function(x, ...) x,
    #                     oob = identity,limits=range(breaks), space="Lab")+
    facet_wrap(~Year, ncol = 1)
    print(paste0("saving plot ", Yr))
    ggsave(filename = paste0("./fig_output/cdec/hgm_snowcdec_",Yr,".png"),
           width = 8, height=8, units = "in",dpi = 150)
}
  
# Step 1: Make Plots For Year Range 
seq(from = 1984, to=2015, by=1) %>% 
  map(cdec_map)

# Step 2: List those Plots, Read them in, and then make animation
list.files(path = "../fig_output/cdec/", pattern = "*.png", full.names = T) %>% 
  map(image_read) %>% # reads each path file
  image_join() %>% # joins image
  image_animate(fps=2.5) %>% # animates, can opt for number of loops
  image_write("../fig_output/cdec_snow.gif") # write to current dir

```

```{r ndvi animation}

mod.hgm<-read_rds(path="../data_output/hgm_gee_mdws_dat.rds") # full model dataset
mod.hgm$month<-lubridate::month(mod.hgm$date)
ndvi.df <- mod.hgm %>% 
  group_by(UCDID, WY, month) %>% 
  filter(index=="NDVI", month==8, mean==max(mean, na.rm=T)) %>%
  select(UCDID, WY, month, mean, median, stdDev, hgm_class_comb, LF:Depressional, LAT_DD, LONG_DD)

#load("../data_output/model_datasets_Aug_max.rda") # data

## Set the NDVI Palette
breaks<-(c(0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8)) # for color scale
palette<-c("darkred","red4","red2","lightcoral","thistle1","limegreen","green2","forestgreen")

# Get the basemap
ca <- get_map(
  #location='california',
  c(lon=-120,lat=38.7),
  zoom=7,crop=T,
  scale="auto",color="bw",source="google",
  maptype="terrain") # can change to terrain

gg <- ggmap(ca, extent='panel',padding=0) # call the basemap, also: extent="panel",fullpage=TRUE

ndvi_map <- function(Yr){
  
  gg + geom_point(data=ndvi.df[ndvi.df$WY==Yr,], 
                        aes(x=LONG_DD, y=LAT_DD, fill=mean),
                  show.legend=F, pch=21, size=4.8, color="gray30")+ 
    theme_bw() + ylab("Latitude") + xlab("Longitude") +
    theme(axis.text.x = element_text(angle = 60, vjust=0.15, size=8),
          legend.position=c(1,1),legend.justification=c(1,1),
          legend.direction="vertical",legend.text=element_text(size=8),
          legend.title=element_text(size=8, face="bold"),
          legend.box="horizontal", panel.background = element_blank(),
          legend.box.just = c("top"), 
          legend.background = element_rect(fill=alpha('white', 0.6), colour = "gray30")) +
    # scale_fill_viridis(name="NDVI (Aug)", limits=range(breaks), 
    #                    breaks=breaks, option = "B", direction = -1)+
    #scale_fill_gradientn("NDVI (Aug)", colours=brewer.pal(8,name="RdYlGn"),
    #                     limits=range(breaks), breaks=breaks)+
    scale_fill_gradientn("NDVI (Aug)",colours=palette,
                         breaks=breaks,values=breaks, rescaler = function(x, ...) x,
                         oob = identity,limits=range(breaks), space="Lab")+
    facet_wrap(~WY, ncol = 1)
    print(paste0("saving plot ", Yr))
    ggsave(filename = paste0("../fig_output/ndvi/hgm_ndvi_",Yr,".png"),
           width = 8,height=8,dpi = 150)
}
  
# Step 1: Make Plots For Year Range 
seq(from = 1984, to=2015, by=1) %>% 
  map(ndvi_map)


# Step 2: List those Plots, Read them in, and then make animation
list.files(path = "../fig_output/ndvi/", pattern = "*.png", full.names = T) %>% 
  map(image_read) %>% # reads each path file
  image_join() %>% # joins image
  image_animate(fps=2) %>% # animates, can opt for number of loops
  image_write("ndvi_aug_hgm.gif") # write to current dir

```

```{r animation w systemImageMagick}

## Using ImageMagick: Set working dir first to images
system(command= "convert ~/Documents/github/meadow_hgm_waterchem/fig_output/cdec/hgm_snowcdec_v2* -delay 100 -loop 0 cdec_snow_IM.gif")
# play with -delay as necessary, this is much faster than the fps in "magick"

# To Resize Use
system(command= "convert -size 1100x850 cdec_snow.gif -resize 800x600 cdec_snow_small.gif")
system(command= "convert cdec_snow.gif -resize 50% cdec_snow_small.gif")

## TO Slow down use existing Gif

system(command = "convert -delay 40 cdec_snow_v2.gif cdec_snow_v3.gif")

## Remove pngs if you want to save space
#file.remove(list.files(pattern=".png"))


```


