---
title: "model Van Norden"
author: "Ryan Peek"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    highlight: pygments
    theme: yeti
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The goal of the modeling is to identify relationships between water source 
(i.e., whether the meadow region in question is driven primarily *snow* or 
*groundwater*) and underlying hydrology/geomorphology. To identify potential
links, we will assess the strength of patterns associated with hydrogeomorphic
types (**HGM**), and water chemistry in conjunction with remote sensing data
from Google Earth Engine (NDVI and NDWI).

## Pilot Meadow: Van Norden

Van Norden is a large meadow at approximately 7,000 ft. 

## Load Required Packages
First step is to load libraries and data required for analysis.

```{r load-libs, include=TRUE}
suppressPackageStartupMessages({
#  library(plyr)
  library(dplyr)
	library(ggplot2)
	library(lubridate)
	library(leaflet)
  library(readr)
  })
```

```{r load-data}

vn.vi<-read_csv("./data/mdws/mdws_batch_NDVI_VanNorden.csv")
dim(vn.vi)
head(vn.vi)


```

