---
title: "Modeling Sourcewater: Van Norden"
author: "Ryan Peek"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  html_document:
    highlight: pygments
    theme: yeti
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Modeling Sourcewater

The goal is to identify relationships between water source 
(i.e., whether the meadow in question in question is driven primarily *snow* or *groundwater*) and underlying hydrology/geomorphology. To identify potential links, we will assess patterns between hydrogeomorphic types (**HGM**), water chemistry, and remote sensing data
from Google Earth Engine (NDVI and NDWI).

# Pilot Meadow: Van Norden

Van Norden is a large meadow at approximately 7,000 ft. 

## Load Required Packages
First step is to load libraries and data required for analysis.

```{r load-libs, include=TRUE}
suppressPackageStartupMessages({
  library(plyr)
  library(dplyr)
	library(ggplot2)
	library(lubridate)
	library(leaflet)
  library(readr)
  library(viridis)
  })
```

```{r load-data}
# read in GEE data
vn.wi<-read_csv("./data/mdws/mdws_batch_NDWI_VanNorden.csv")
vn.vi<-read_csv("./data/mdws/mdws_batch_NDVI_VanNorden.csv")

source(file='scripts/functions/gee_data_cleanup.R')
ndwi_df <- gee_tidy_df(vn.wi) # clean up df
ndvi_df <- gee_tidy_df(vn.vi)
vn.dat<-bind_rows(ndwi_df,ndvi_df) # bind both df together
# rm(vn.wi, vn.vi, ndvi_df, ndwi_df) # remove temporary data

source("./scripts/functions/doy.R")
head(vn.dat)
vn.dat$date<-ymd(vn.dat$date)
vn.dat$WY<-wtr_yr(vn.dat$date)
vn.dat$DOWY<-dowy(vn.dat$date)
vn.dat$DOY<-yday(vn.dat$date)
#saveRDS(vn.dat, file="data_output/VanNorden_GEE.rds") # save data 

# read in shape data
library(foreign)
mdw_dbf <- read.dbf("data/shps/Sierra_Nevada_MultiSource_Meadow_Polygons_Compilation_v1_simplify10m.dbf")

# filter to Van Norden
meadow <- 'Van Norden'
mdwID <- 'UCDSNM014503'
mdw_vn <- mdw_dbf %>% filter(ID == mdwID)

```

### Variability During Drought Years

Plots of 3 years of data, 2011 was a wet year, followed by drought. Notice the pattern of lower peaks and an earlier maximum, particularly noticeable in the NDWI data.

```{r plot-ndvi-variability, include=TRUE, eval=TRUE, echo=FALSE}

source("./scripts/multiplot.R")

y1 <- 2011
meadow <- 'Van Norden'
mdwID <- 'UCDSNM014503'

single_y1 <- vn.dat[vn.dat$WY==y1 & vn.dat$ID==mdwID & vn.dat$index=="NDVI",]

p1<- ggplot(data=single_y1, aes(x=DOWY, y=mean)) + 
	geom_errorbar(aes(ymax = mean + stdDev, ymin=mean - stdDev), width=0.2, color="gray30", alpha=0.5) +
  stat_smooth(color="#440154FF", alpha=0.8)+
  geom_point(fill="#440154FF", pch=21) + # add points
	scale_y_continuous( limits = c(-0.5,1), expand = c(0,0) ) +
	theme_bw() + ylab("NDVI") + xlab("Day of Water Year") +
	ggtitle(paste0("NDVI with SD within Van Norden Meadow for ", y1))

y2 <- 2013
single_y2 <- vn.dat[vn.dat$WY==y2 & vn.dat$ID==mdwID & vn.dat$index=="NDVI",]

p2<-ggplot(data=single_y2, aes(x=DOWY, y=mean)) + 
	geom_errorbar(aes(ymax = mean + stdDev, ymin=mean - stdDev), width=0.2, alpha=0.8, color="gray30") +
  stat_smooth(color="#21908CFF", alpha=0.5)+
  geom_point(fill="#21908CFF", pch=21)+ # add points
  scale_y_continuous( limits = c(-0.5,1), expand = c(0,0) ) +
	theme_bw() + # set to bw theme
	ylab("NDVI") + xlab("Day of Water Year") + # Set axis labels
	ggtitle(paste0("NDVI with SD within Van Norden Meadow for ", y2))  # title of graph


y3 <- 2015
single_y3 <- vn.dat[vn.dat$WY==y3 & vn.dat$ID==mdwID & vn.dat$index=="NDVI",]

p3<-ggplot(data=single_y3, aes(x=DOWY, y=mean)) + 
	geom_errorbar(aes(ymax = mean + stdDev, ymin=mean - stdDev), width=0.2, alpha=0.8, color="gray30") +
  stat_smooth(color="#35B779FF", alpha=0.5) +
  geom_point(fill="#35B779FF", pch=21)+ # add points
  scale_y_continuous( limits = c(-0.5,1), expand = c(0,0) ) +
	theme_bw() + # set to bw theme
	ylab("NDVI") + xlab("Day of Water Year") + # Set axis labels
	ggtitle(paste0("NDVI with SD within Van Norden Meadow for ", y3))  #

multiplot(p1, p2, p3) 

#ggsave(multiplot(p2011, p2014), filename = "fig_output/VN_2011-2014_NDVI.svg",
#       width = 9, height = 6, units = "in", dpi = 200) # make sure 'svglite' installed
```

NDVI is vegetation greeness, so differences are more pronounced during the growing season. In 2011, there was more snow, which remained on the meadow longer and the growing season started later, so values are actually lower earlier in the year. In 2014 the opposite was true, due to a very dry water year which actually permitted an earlier start to vegetation growth in the meadow.

```{r plot-ndwi-variability, include=TRUE, eval=TRUE, echo=FALSE}

source("./scripts/multiplot.R")

y1 <- 2011
meadow <- 'Van Norden'
mdwID <- 'UCDSNM014503'

single_y1 <- vn.dat[vn.dat$WY==y1 & vn.dat$ID==mdwID & vn.dat$index=="NDWI",]

p1<- ggplot(data=single_y1, aes(x=DOWY, y=mean)) + 
	geom_errorbar(aes(ymax = mean + stdDev, ymin=mean - stdDev), width=0.2, color="gray30", alpha=0.5) +
  stat_smooth(color="#440154FF", alpha=0.2)+
  geom_point(fill="#440154FF", pch=21) + # add points
	scale_y_continuous( limits = c(-0.5,1), expand = c(0,0) ) +
	theme_bw() + ylab("NDWI") + xlab("Day of Water Year") +
	ggtitle(paste0("NDWI with SD within Van Norden Meadow for ", y1))

y2 <- 2013
single_y2 <- vn.dat[vn.dat$WY==y2 & vn.dat$ID==mdwID & vn.dat$index=="NDWI",]

p2<-ggplot(data=single_y2, aes(x=DOWY, y=mean)) + 
	geom_errorbar(aes(ymax = mean + stdDev, ymin=mean - stdDev), width=0.2, alpha=0.8, color="gray30") +
  stat_smooth(color="#21908CFF", alpha=0.2)+
  geom_point(fill="#21908CFF", pch=21)+ # add points
  scale_y_continuous( limits = c(-0.5,1), expand = c(0,0) ) +
	theme_bw() + # set to bw theme
	ylab("NDVI") + xlab("Day of Water Year") + # Set axis labels
	ggtitle(paste0("NDWI with SD within Van Norden Meadow for ", y2))  # title of graph


y3 <- 2015
single_y3 <- vn.dat[vn.dat$WY==y3 & vn.dat$ID==mdwID & vn.dat$index=="NDWI",]

p3<-ggplot(data=single_y3, aes(x=DOWY, y=mean)) + 
	geom_errorbar(aes(ymax = mean + stdDev, ymin=mean - stdDev), width=0.2, alpha=0.8, color="gray30") +
  stat_smooth(color="#35B779FF", alpha=0.2)+
  geom_point(fill="#35B779FF", pch=21)+ # add points
  scale_y_continuous( limits = c(-0.5,1), expand = c(0,0) ) +
	theme_bw() + # set to bw theme
	ylab("NDVI") + xlab("Day of Water Year") + # Set axis labels
	ggtitle(paste0("NDWI with SD within Van Norden Meadow for ", y3))  #

multiplot(p1, p2, p3) 

```

Now the plot should show some differences between NDVI in a wet year and a dry year
in Van Norden Meadow. What we need to do next is parse this out into proportion 
of HGM types, and build a model which will give us a distribution of probabilities 
associated with a given HGM type vs. NDVI or NDWI, during a given growing season.


## Add a Shiny App

# SHINY PLOT --------------------------------------------------------------
```{r}

library(shiny)
shinyApp(
  ui = fluidPage(
    h3("NDVI-NDWI"),
    inputPanel(
      selectInput("data", label="Data Interval",
                  choices=c("WY","DOWY","DOY","date"), 
                  selected="WY"),
      selectInput("datatype", label = "Mean/Median/SD", 
                  choices=c("mean","median","stdDev")),
      radioButtons("type", 
                   label = "NDVI or NDWI", 
                   choices = c("NDVI","NDWI"),selected = "NDVI"),
      checkboxInput("years", label = "Pick one Year", value=FALSE),
        conditionalPanel(
          condition="input.years==true",
          sliderInput("wtyr",value = 1985, 
                  min = 1984, max = 2015, 
                  step = 1, sep = FALSE,
                  label="Pick a Year"))
    ),
    mainPanel(
      plotOutput("Plot", width = "800px", height="500px")
    )
  ),
  
  server = function(input, output) {
    
    output$Plot<-renderPlot({
      
      df<-reactive({
        if(input$years == TRUE) {
          dplyr::filter(vn.dat, index==input$type, WY %in% input$wtyr)}
        else{
          dplyr::filter(vn.dat, index==input$type)
        }
      })
      
      print(ggplot() + 
              geom_smooth(
                data=vn.dat[vn.dat$index==input$type & vn.dat$WY %in% input$wtyr,], aes_string(x=input$data, y=input$datatype),
                          color="#21908CFF")+
        geom_point(data=vn.dat[vn.dat$index==input$type & vn.dat$WY %in% input$wtyr,], 
                   aes_string(x=input$data, y=input$datatype),
                   color="#21908CFF", pch=21) + 
        theme_bw()
      )
    })
  }
)
```

## Time for a Model

Use multilevel but use year as a inflation stat/metric?

